---
title: "Class 1 Survey Analysis"
author: "Supriya"
date: "2024-09-14"
output: html_document
---
### **INSTRUCTIONS**
####  **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks. Compare and consolidate your answers. Select one member of your group to walk the class through how each problem was solved. Elect one member of your group to turn in your consolidated code with answers.**

##
##

##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA-2024) (The dataset is called 'Class 1 Survey Fall 2024_di.csv'), calling the R dataframe that you create *C1survey*.
```{r}

#Clearing environmental variables:
rm(list = ls())

# Loading packages dplyr and readr
pacman::p_load(dplyr, readr)

# Importing the C1survey dataset from Github
C1survey<-read_csv("https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv")
```

##
##

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.    
```{r}
# Determining the number of observations and variables in C1survey dataset
dim(C1survey)
# there are 29 observations and 27 variables

#Determining the structure of C1survey
str(C1survey)
```

##
##

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed.
```{r}
# a. Using names function to rename the 27 variables in order

names(C1survey)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city", "highest_educ_level")

# b. Using colnames function to display the renamed variables
colnames(C1survey)
```
##
##

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset.
```{r}

#Using table sapply function to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset

table(sapply(C1survey, class)) 

# There are 23 character and 4 integer

summary(C1survey)
```

##
##
##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values. If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
# a. Checking for any unusual or missing values for *bday* and *bmonth* variables

C1survey$bday
C1survey$bmonth

# b. recode bday with non numeric values

# Recoding *bday*
class(C1survey$bday) 
C1survey$bday<-ifelse(C1survey$bday == "May 21-report 21", "21", C1survey$bday)
C1survey$bday<-ifelse(C1survey$bday == "March 31st", "31", C1survey$bday)
C1survey$bday 

# Checking values
class(C1survey$bday) 

# Converting *bday* variable to numeric
C1survey$bday<-as.numeric(C1survey$bday)
table(C1survey$bday, useNA = "always") # check values 

# bmonth recoding
class(C1survey$bmonth)

C1survey$bmonth[C1survey$bmonth=="May 21-report 5"]<-"5"
C1survey$bmonth[C1survey$bmonth=="February"]<-"2"
C1survey$bmonth[C1survey$bmonth=="September"]<-"9"
C1survey$bmonth[C1survey$bmonth=="July"]<-"7"
C1survey$bmonth[C1survey$bmonth=="March"]<-"3"
C1survey$bmonth # check values 

# Converting *bmonth* variable to numeric
C1survey$bmonth<-as.numeric(C1survey$bmonth)

# check values
C1survey$bmonth 

# c. calculating medians for bday and bmonth

median(C1survey$bday, na.rm=TRUE) 

# The median birth day is 14. However, this result is biased due to presence of 0, which is not a value.

median(C1survey$bmonth, na.rm=TRUE) 

# The median birth month is 7 i.e. July.
```
##
##
##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 
```{r}
# a. Creating a new variable *bseason*

C1survey <- C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(C1survey$bmonth, C1survey$bseason) 

# checking variable

C1survey$bseason_f<-factor(C1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))

# b. Using the table function to print a table of Seasons in the columns and bmonths in the rows

table(C1survey$bmonth, C1survey$bseason)

#Checking values

table(C1survey$bmonth, C1survey$bseason_f)

# c.
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason_f, useNA = "always"), 1)
tab

# Summing the columns using the addmargins function 

#Births by season: Fall = 4 Winter = 9 Spring = 8 Summer = 8
## Majority of the students (9 out of 29 students) were born in winter, followed by 8 students born in Spring and Summer each, and only 4 students were born in fall.
```
##
##
##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).

```{r}

# Question: What percentage of students like the season that they were born in the most?
## Variables to analyze: student's favorite seasons (fav_season) and seasons that a student was born in (bseason)

# Ensuring that the fav_season and bseason variables are categorical variables
C1survey$fav_season <- as.factor(C1survey$fav_season)
C1survey$bseason <- as.factor(C1survey$bseason)

# Creating a new variable 'likes_born_season' that has a value '1' if the favorite season matches the birth season, and '0' if they don't match.
C1survey <- C1survey %>%
  mutate(likes_born_season = ifelse(fav_season == bseason, 1, 0))

# Calculating the percentage of students who like the season they were born in
percentage_likes_born_season <- mean(C1survey$likes_born_season, na.rm = TRUE) * 100

# Display the result and rounding off the percentage to two decimal points
paste("Percentage of students who like the season they were born in:", round(percentage_likes_born_season, 2), "%")

#20.69% of the students liked the seasons that they were born in the most.
```

